|

### 2025-10-20 manuscript/manuscript_preprint.qmd
update mylibrary.bib file, update eqs in manuscript, forgot to commit message previous intro add

---
title: "Methods and Results"
format: docx
bibliography: 
  mylibrary.bib
csl: apa.csl
execute: 
  echo: false
  message: false
  warning: false
---
```{r}
library(tidyverse)
library(flextable)
library(tidybayes)

@@

# Methods
This systematic review and meta-analysis was pre-registered on PROSPERO ([CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) initially on the $3^{rd}$ of December 2024. The primary aim of this review was to examine the descriptive question "Does resting energy expenditure (REE) differ between women with and without polycystic ovary syndrome (PCOS)?".

## Search Strategy
PubMed, Web of Science, and MEDLINE databases were searched using the following Boolean search string: (("Basal Metabolic Rate"[MeSH] OR "Energy Metabolism"[MeSH] OR "Resting Metabolic Rate" OR RMR OR "Resting Energy Expenditure" OR REE OR "Basal Metabolic Rate" OR BMR OR "resting energy" OR "basal energy expenditure") AND ("Polycystic Ovary Syndrome"[MeSH] OR "Polycystic Ovary Syndrome" OR PCOS OR "Polycystic Ovarian Disease" OR "Stein-Leventhal Syndrome")). Searches were limited to publications up until May 2025 when the search was completed, limited to English language articles, and Rayyan was used to manage the search and screening process.

@@

<!-- ## Risk of bias (quality) assessment -->
<!-- The assessment will be performed at the study level with the Joanna Briggs Institute (JBI) Critical Appraisal Checklist for Analytical Cross Sectional Studies, which assesses the methodological quality of a study and determines the extent to which a study has addressed the possibility of bias in its design, conduct and analysis. -->

@@

The main model for mean REE with $\textrm{cond}$ representing the condition (either control or PCOS) was as follows:
<!-- $$ -->
<!-- \begin{align*} -->
<!-- \hat{\theta}_{ijkl} &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt] -->

<!-- \mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_[ijkl] -->
<!--        + \alpha_{0,\mathrm{lab}[i]} + \alpha_{1,\mathrm{lab}[i]}\,\text{cond} -->
<!--        + \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond} -->
<!--        + \alpha_{0,\mathrm{arm}[k]} -->
<!--        + \alpha_{0,\mathrm{effect}[l]} \\[10pt] -->

<!-- \begin{pmatrix} -->
<!-- \alpha_{0,\mathrm{lab}[i]} \\[4pt] -->
<!-- \alpha_{1,\mathrm{lab}[i]} -->
<!-- \end{pmatrix} -->
<!-- &\sim  -->
<!-- \mathcal{N}\!\left( -->
<!-- \begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix}, -->
<!-- \Sigma_{\mathrm{lab}} -->
<!-- \right), -->
<!-- \quad -->
<!-- \Sigma_{\mathrm{lab}} = -->
<!-- \begin{pmatrix} -->
<!-- \sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} \\[4pt] -->
<!-- \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}} -->
<!-- \end{pmatrix}  -->

<!-- \text{, for lab i = 1,} \dots \text{,I} \\[12pt] -->

<!-- \begin{pmatrix} -->
<!-- \alpha_{0,\mathrm{study}[j]} \\[4pt] -->
<!-- \alpha_{1,\mathrm{study}[j]} -->
<!-- \end{pmatrix} -->
<!-- &\sim  -->
<!-- \mathcal{N}\!\left( -->
<!-- \begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix}, -->
<!-- \Sigma_{\mathrm{study}} -->
<!-- \right), -->
<!-- \quad -->
<!-- \Sigma_{\mathrm{study}} = -->
<!-- \begin{pmatrix} -->
<!-- \sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} \\[4pt] -->
<!-- \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}} -->
<!-- \end{pmatrix}  -->
<!-- \text{, for study j = 1,} \dots \text{,J} \\[12pt] -->

<!-- \alpha_{0,\mathrm{arm}[k]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}})  -->
<!-- \text{, for arm k = 1,} \dots \text{,K} \\[12pt] -->
<!-- \alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}})  -->
<!-- \text{, for effect l = 1,} \dots \text{,L} \\[12pt] -->

<!-- \end{align*} -->
<!-- $$ -->

$$
\hat{\theta}_{ijkl} \sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl})
$$

$$
\mu_{ijkl} = \beta_0 + \beta_1\,\text{cond}_{ijkl} 
+ \alpha_{0,\mathrm{lab}[i]} + \alpha_{1,\mathrm{lab}[i]}\,\text{cond} 
+ \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond} 
+ \alpha_{0,\mathrm{arm}[k]} + \alpha_{0,\mathrm{effect}[l]}
$$

$$
\begin{pmatrix}
\alpha_{0,\mathrm{lab}[i]} \\
\alpha_{1,\mathrm{lab}[i]}
\end{pmatrix}
\sim 
\mathcal{N}\Bigg(
\begin{pmatrix} 0 \\ 0 \end{pmatrix},
\Sigma_{\mathrm{lab}}
\Bigg), 
\quad
\Sigma_{\mathrm{lab}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}} \sigma_{0,\mathrm{lab}} \sigma_{1,\mathrm{lab}} \\
\rho_{\mathrm{lab}} \sigma_{0,\mathrm{lab}} \sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}}
\end{pmatrix},
\text{ for lab } i = 1, \dots, I
$$

$$
\begin{pmatrix}
\alpha_{0,\mathrm{study}[j]} \\
\alpha_{1,\mathrm{study}[j]}
\end{pmatrix}
\sim 
\mathcal{N}\Bigg(
\begin{pmatrix} 0 \\ 0 \end{pmatrix},
\Sigma_{\mathrm{study}}
\Bigg), 
\quad
\Sigma_{\mathrm{study}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}} \sigma_{0,\mathrm{study}} \sigma_{1,\mathrm{study}} \\
\rho_{\mathrm{study}} \sigma_{0,\mathrm{study}} \sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}}
\end{pmatrix},
\text{ for study } j = 1, \dots, J
$$

$$
\alpha_{0,\mathrm{arm}[k]} \sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}}), 
\text{ for arm } k = 1, \dots, K
$$

$$
\alpha_{0,\mathrm{effect}[l]} \sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}}), 
\text{ for effect } l = 1, \dots, L
$$

@@

<!-- $$ -->
<!-- \begin{align*} -->
<!-- \beta_0 &\sim \text{Student-}t(3,\, 1441.81,\, 84.56) \\[4pt] -->
<!-- \beta_1 &\sim \text{Student-}t(3,\, 0,\, 200) \\[12pt] -->

<!-- \sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 149.89,\, 82.91) \\[4pt] -->
<!-- \sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\,  -->
<!-- \sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 112.4) \\[12pt] -->

<!-- \mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt] -->
<!-- \mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1) -->
<!-- \end{align*} -->
<!-- $$ -->

$$
\beta_0 \sim \text{Student-}t(3, 1441.81, 84.56) \\
\beta_1 \sim \text{Student-}t(3, 0, 200)
$$

$$
\sigma^2_{0,\mathrm{study}} \sim \text{Half-student-}t(3, 149.89, 82.91) \\
\sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\, 
\sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}},\, \sigma^2_{0,\mathrm{effect}} \sim \text{Half-student-}t(3, 0, 112.4)
$$

$$
\mathrm{corr}_{\mathrm{lab}} \sim \mathrm{LKJ}(1) \\
\mathrm{corr}_{\mathrm{study}} \sim \mathrm{LKJ}(1)
$$

where the prior for $\beta_{0}$, which corresponded to the model intercept and mean REE in the control condition, was set based on meta-analysis of the mean REEs for women from two large studies of healthy people [@velasquezUseAmmoniaInhalants2011; @pavlidouRevisedHarrisBenedict2023] though set with a conservative degrees of freedom for the $\textrm{Student-}t$ distribution.The random intercept $\sigma^2_{0,study}$ was set similarly to this. The prior for the fixed effect $\beta_{1}$, reflecting the difference between control and PCOS conditions was set based on a wide range of possible values considering the minimum and and maximum values of the ranges reported in the two studies noted (i.e., 2492 - 908 = 1584). We then set a prior that permits values approximately across this range of values with the majority of it's mass centred around zero. The remaining random effects were set based on the default weakly regularising priors for `brms` and scaled to the expected response values using a $\textrm{half-student-}t$ distribution with 3 degrees of freedom and $\mu=0$, and both correlation matrices $\textrm{corr}_{lab}$ and $\textrm{corr}_{study}$ were set with an $\textrm{LKJcorr(1)}$ distribution.

@@

<!-- $$ -->
<!-- \begin{align*} -->
<!-- \log(\hat{\theta}_{ijkl}) &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt] -->

<!-- \mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_{[ijkl]}  -->
<!--        + \beta_2\,\tilde{m}_{[ijkl]} -->
<!--        + \alpha_{0,\mathrm{lab}[i]} + \alpha_{1,\mathrm{lab}[i]}\,\text{cond} -->
<!--        + \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond} -->
<!--        + \alpha_{0,\mathrm{arm}[k]} -->
<!--        + \alpha_{0,\mathrm{effect}[l]} \\[8pt] -->

<!-- m &= \log(y_{i,\mathrm{mean,[ijkl]}}) \\[3pt] -->
<!-- \tilde{m} &\sim \mathcal{N}(m,\, \sigma_{\log(y_{i,\mathrm{mean,[ijkl]}})}^2) \\[10pt] -->

<!-- \begin{pmatrix} -->
<!-- \alpha_{0,\mathrm{lab}[i]} \\[4pt] -->
<!-- \alpha_{1,\mathrm{lab}[i]} -->
<!-- \end{pmatrix} -->
<!-- &\sim  -->
<!-- \mathcal{N}\!\left( -->
<!-- \begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix}, -->
<!-- \Sigma_{\mathrm{lab}} -->
<!-- \right), -->
<!-- \quad -->
<!-- \Sigma_{\mathrm{lab}} = -->
<!-- \begin{pmatrix} -->
<!-- \sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} \\[4pt] -->
<!-- \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}} -->
<!-- \end{pmatrix}  -->

<!-- \text{, for lab i = 1,} \dots \text{,I} \\[12pt] -->

<!-- \begin{pmatrix} -->
<!-- \alpha_{0,\mathrm{study}[j]} \\[4pt] -->
<!-- \alpha_{1,\mathrm{study}[j]} -->
<!-- \end{pmatrix} -->
<!-- &\sim  -->
<!-- \mathcal{N}\!\left( -->
<!-- \begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix}, -->
<!-- \Sigma_{\mathrm{study}} -->
<!-- \right), -->
<!-- \quad -->
<!-- \Sigma_{\mathrm{study}} = -->
<!-- \begin{pmatrix} -->
<!-- \sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} \\[4pt] -->
<!-- \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}} -->
<!-- \end{pmatrix}  -->
<!-- \text{, for study j = 1,} \dots \text{,J} \\[12pt] -->

<!-- \alpha_{0,\mathrm{arm}[k]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}})  -->
<!-- \text{, for arm k = 1,} \dots \text{,K} \\[12pt] -->
<!-- \alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}})  -->
<!-- \text{, for effect l = 1,} \dots \text{,L} \\[12pt] -->
<!-- \end{align*} -->
<!-- $$ -->

$$
\log(\hat{\theta}_{ijkl}) \sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl})
$$

$$
\mu_{ijkl} = \beta_0 + \beta_1\,\text{cond}_{[ijkl]} 
+ \beta_2\,\tilde{m}_{[ijkl]}
+ \alpha_{0,\mathrm{lab}[i]} + \alpha_{1,\mathrm{lab}[i]}\,\text{cond} 
+ \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond} 
+ \alpha_{0,\mathrm{arm}[k]} + \alpha_{0,\mathrm{effect}[l]}
$$

$$
m = \log(y_{i,\mathrm{mean,[ijkl]}})
$$

$$
\tilde{m} \sim \mathcal{N}(m, \sigma_{\log(y_{i,\mathrm{mean,[ijkl]}})}^2)
$$

$$
\begin{pmatrix}
\alpha_{0,\mathrm{lab}[i]} \\
\alpha_{1,\mathrm{lab}[i]}
\end{pmatrix}
\sim 
\mathcal{N}\Bigg(
\begin{pmatrix} 0 \\ 0 \end{pmatrix},
\Sigma_{\mathrm{lab}}
\Bigg),
\quad
\Sigma_{\mathrm{lab}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}} \sigma_{0,\mathrm{lab}} \sigma_{1,\mathrm{lab}} \\
\rho_{\mathrm{lab}} \sigma_{0,\mathrm{lab}} \sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}}
\end{pmatrix},
\text{ for lab } i = 1, \dots, I
$$

$$
\begin{pmatrix}
\alpha_{0,\mathrm{study}[j]} \\
\alpha_{1,\mathrm{study}[j]}
\end{pmatrix}
\sim 
\mathcal{N}\Bigg(
\begin{pmatrix} 0 \\ 0 \end{pmatrix},
\Sigma_{\mathrm{study}}
\Bigg),
\quad
\Sigma_{\mathrm{study}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}} \sigma_{0,\mathrm{study}} \sigma_{1,\mathrm{study}} \\
\rho_{\mathrm{study}} \sigma_{0,\mathrm{study}} \sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}}
\end{pmatrix},
\text{ for study } j = 1, \dots, J
$$

$$
\alpha_{0,\mathrm{arm}[k]} \sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}}), 
\text{ for arm } k = 1, \dots, K
$$

$$
\alpha_{0,\mathrm{effect}[l]} \sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}}), 
\text{ for effect } l = 1, \dots, L
$$

@@

The priors for this model were as follows:
<!-- $$ -->
<!-- \begin{align*} -->
<!-- \beta_0 &\sim \text{Student-}t(3,\, 5.54,\, 0.80) \\[4pt] -->
<!-- \beta_1 &\sim \text{Student-}t(3,\, 0,\, 5.3) \\[4pt] -->
<!-- \beta_2 &\sim \text{Student-}t(3,\, 0,\, 2.5) \\[12pt] -->

<!-- \text{mean}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 7.28,\, 0.62) \\[4pt] -->
<!-- \text{sd}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 0,\, 5) \\[12pt] -->

<!-- \sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 1.08,\, 1.06) \\[4pt] -->
<!-- \sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\, -->
<!-- \sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 2.5) \\[12pt] -->

<!-- \mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt] -->
<!-- \mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1) -->
<!-- \end{align*} -->

<!-- $$ -->

$$
\beta_0 \sim \text{Student-}t(3, 5.54, 0.80) \\
\beta_1 \sim \text{Student-}t(3, 0, 5.3) \\
\beta_2 \sim \text{Student-}t(3, 0, 2.5)
$$

$$
\text{mean}(\tilde{m}) \sim \text{Half-student-}t(3, 7.28, 0.62) \\
\text{sd}(\tilde{m}) \sim \text{Half-student-}t(3, 0, 5)
$$

>>>>

---
title: |
  Does resting energy expenditure differ between women with and without polycystic ovary syndrome (PCOS): a systematic review and meta-analysis
format:
  pdf: default
  aft-pdf:
    keep-tex: true
author:
  - name: Richard Kirwan
  - name: Leigh Peele
  - name: Greg Nuckols
  - name: Georgia Kohlhoff
  - name: Hannah Cabré
  - name: Alyssa Olenick
  - name: James Steele
abstract: |
  Background: Polycystic ovary syndrome is the most common endocrine condition affecting women of reproductive age. Overweight and obesity in women with PCOS is commonly deemed a result of a lower basal metabolic rate (BMR), which has influenced lifestyle-based interventions in this group. Evidence supporting this reduced, however, is not consistent. Objectives: To systematically search and meta-analyse the existing literature to determine if resting energy expenditure differs between women with and without polycystic ovary syndrome (PCOS). Design: A systematic search was conducted using PubMed, Medline and Web of Science databases from January 1990 to January 2025. Studies that assessed resting energy expenditure in women with and without PCOS were included. Results: 18 studies were identified. Study populations ranged from X to X participants, with mean ages ranging from X to X years. Conclusions: TO ADD
keywords: Polycystic ovary syndrome, PCOS, resting energy expenditure, REE, basal metabolic rate, BMR
pdf-engine: xelatex
bibliography: 
  mylibrary.bib
csl: apa.csl
execute: 
  echo: false
  message: false
  warning: false
---
```{r}
library(tidyverse)
library(flextable)
library(tidybayes)

@@

# Introduction
Polycystic ovary syndrome (PCOS) affects approximately 10% of women of reproductive age worldwide, making it the most common endocrine disorder affecting this population [@salariGlobalPrevalencePolycystic2024]. Due to a number of factors including hyperandrogenism and alterations in insulin resistance, PCOS is believed to contribute to an increased risk of diabetes, metabolic syndrome and cardiovascular disease [@glintborgCardiovascularDiseaseNationwide2018; @glintborgProspectiveRiskType2022; @perssonHigherRiskType2021; @glintborgProspectiveRiskType2024; @limMetabolicSyndromePolycystic2019], along with being a leading cause of anovulatory infertility in women [@johamPrevalenceInfertilityUse2015]. Furthermore, epidemiological data consistently demonstrates that women with PCOS are significantly more likely to suffer from overweight or obesity, compared to the general female population, with estimates ranging from 38% to 88% of PCOS patients falling into these BMI categories [@barberWhyAreWomen2022; @barberObesityPolycysticOvary2021].

The elevated incidence of overweight and obesity in PCOS is likely multifactorial with proposed mechanisms including blunted postprandial appetite hormone responses leading to reduced satiety and increased food cravings [@hirschbergImpairedCholecystokininSecretion2004; @moranGhrelinMeasuresSatiety2004; @stefanakiFoodCravingsObesity2024] and a reduced basal metabolic rate (BMR) [@georgopoulosBasalMetabolicRate2009]. Indeed, the relevance of this reduced BMR, has become widely accepted amongst the scientific and medical community, with the original publication which reported this from Georgopoulos et al. [-@georgopoulosBasalMetabolicRate2009] having 109 citations according to Google Scholar metrics, as of writing (Add date and link!!!!). Furthermore, this publication has been widely disseminated amongst the lay community interested in PCOS, particularly in relation to lifestyle interventions, featuring in the top seven Google search results as of writing. However, despite the wide acceptance of this reported reduced BMR in PCOS, multiple other publications have found no such difference between women with and without PCOS [@graffSaturatedFatIntake2017; @larssonDietaryIntakeResting2016; @segalRestingMetabolicRate1990].

The consequences of this wide acceptance of a reduced BMR in women with PCOS should not be underestimated, particularly in light of the aforementioned incidence of overweight and obesity in this population. It has been well documented that women with PCOS suffer with anxiety, depression and a lower quality of life (QOL) as a result of negative body image due to, amongst other reasons, body weight concerns [@himeleinDepressionBodyImage2006; @hofmannBodyImageMental2025; @gellerBodyImageIllness2025; @davitadzeBodyImageConcerns2023]. Consequently, women with PCOS have been observed to engage in more frequent weight-loss attempts than those without [@pesonenPolycysticOvarySyndrome2023]. Furthermore, intentional weight-loss is recognised as efficacious for improving the symptoms of PCOS [@aleneziImpactDietInducedWeight2024; @holteRestoredInsulinSensitivity1995] and, as such, is often recommended [@teedeRecommendations2023International2023]. However, while weight-loss is frequently pursued for body image and/or health-related reasons, women with PCOS may be discouraged from doing so due to the commonly held belief of lower BMR making weight loss attempts more difficult.

Therefore, to determine if there is indeed a difference in resting energy expenditure (REE) between women with and without PCOS, we completed a systematic review and meta-analysis of studies reporting REE in these populations.

# Methods
This systematic review and meta-analysis was pre-registered on PROSPERO ([CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) initially on the $3^{rd}$ of December 2024 and performed in accordance with the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) statement guidelines [@liberatiPRISMAStatementReporting2009]. The primary aim of this review was to examine the descriptive question "Does resting energy expenditure (REE) differ between women with and without polycystic ovary syndrome (PCOS)?".

## Search Strategy
PubMed, Web of Science, and MEDLINE databases were searched using the following Boolean search string: (("Basal Metabolic Rate"[MeSH] OR "Energy Metabolism"[MeSH] OR "Resting Metabolic Rate" OR RMR OR "Resting Energy Expenditure" OR REE OR "Basal Metabolic Rate" OR BMR OR "resting energy" OR "basal energy expenditure") AND ("Polycystic Ovary Syndrome"[MeSH] OR "Polycystic Ovary Syndrome" OR PCOS OR "Polycystic Ovarian Disease" OR "Stein-Leventhal Syndrome")). Searches were limited to publications up until May 2025 when the search was completed, limited to English language articles, and Rayyan was used to manage the search and screening process. Two reviewers (RK and GK) independently screened all titles and abstracts against the predefined inclusion and exclusion criteria. Articles deemed potentially eligible by either reviewer were retrieved in full text. Full texts were then independently assessed by RK and LP to determine final eligibility. Any disagreements at either stage were resolved through discussion, and when consensus could not be reached, a third reviewer acted as an adjudicator.

@@

@@

The main model for mean REE with $\textrm{cond}$ representing the condition (either control or PCOS) was as follows:
$$
\small
\begin{align*}
\hat{\theta}_{ijkl} &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt]

\mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_[ijkl]
       + \alpha_{0,\mathrm{lab}[i]} + \alpha_{1,\mathrm{lab}[i]}\,\text{cond}
       + \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond}
       + \alpha_{0,\mathrm{arm}[k]}
       + \alpha_{0,\mathrm{effect}[l]} \\[10pt]

\begin{pmatrix}
\alpha_{0,\mathrm{lab}[i]} \\[4pt]
\alpha_{1,\mathrm{lab}[i]}
\end{pmatrix}
&\sim
\mathcal{N}\!\left(
\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
\Sigma_{\mathrm{lab}}
\right),
\quad
\Sigma_{\mathrm{lab}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} \\[4pt]
\rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}}
\end{pmatrix}

\text{, for lab i = 1,} \dots \text{,I} \\[12pt]

\begin{pmatrix}
\alpha_{0,\mathrm{study}[j]} \\[4pt]
\alpha_{1,\mathrm{study}[j]}
\end{pmatrix}
&\sim
\mathcal{N}\!\left(
\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
\Sigma_{\mathrm{study}}
\right),
\quad
\Sigma_{\mathrm{study}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} \\[4pt]
\rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}}
\end{pmatrix}
\text{, for study j = 1,} \dots \text{,J} \\[12pt]

\alpha_{0,\mathrm{arm}[k]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}})
\text{, for arm k = 1,} \dots \text{,K} \\[12pt]
\alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}})
\text{, for effect l = 1,} \dots \text{,L} \\[12pt]

\end{align*}
$$

@@

$$
\small
\begin{align*}
\beta_0 &\sim \text{Student-}t(3,\, 1441.81,\, 84.56) \\[4pt]
\beta_1 &\sim \text{Student-}t(3,\, 0,\, 200) \\[12pt]

\sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 149.89,\, 82.91) \\[4pt]
\sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\,
\sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 112.4) \\[12pt]

\mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt]
\mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1)
\end{align*}
$$
where the prior for $\beta_{0}$, which corresponded to the model intercept and mean REE in the control condition, was set based on meta-analysis of the mean REEs for women from two large studies of healthy people [@velasquezUseAmmoniaInhalants2011; @pavlidouRevisedHarrisBenedict2023] though set with a conservative degrees of freedom for the $\textrm{Student-}t$ distribution.The random intercept $\sigma^2_{0,study}$ was set similarly to this. The prior for the fixed effect $\beta_{1}$, reflecting the difference between control and PCOS conditions was set based on a wide range of possible values considering the minimum and and maximum values of the ranges reported in the two studies noted (i.e., 2492 - 908 = 1584). We then set a prior that permits values approximately across this range of values with the majority of it's mass centred around zero. The remaining random effects were set based on the default weakly regularising priors for `brms` and scaled to the expected response values using a $\textrm{half-student-}t$ distribution with 3 degrees of freedom and $\mu=0$, and both correlation matrices $\textrm{corr}_{lab}$ and $\textrm{corr}_{study}$ were set with an $\textrm{LKJcorr(1)}$ distribution.

@@

$$
\small
\begin{align*}
\log(\hat{\theta}_{ijkl}) &\sim \mathcal{N}(\mu_{ijkl}, \sigma^2_{ijkl}) \\[6pt]

\mu_{ijkl} &= \beta_0 + \beta_1\,\text{cond}_{[ijkl]}
       + \beta_2\,\tilde{m}_{[ijkl]}
       + \alpha_{0,\mathrm{lab}[i]} + \alpha_{1,\mathrm{lab}[i]}\,\text{cond}
       + \alpha_{0,\mathrm{study}[j]} + \alpha_{1,\mathrm{study}[j]}\,\text{cond}
       + \alpha_{0,\mathrm{arm}[k]}
       + \alpha_{0,\mathrm{effect}[l]} \\[8pt]

m &= \log(y_{i,\mathrm{mean,[ijkl]}}) \\[3pt]
\tilde{m} &\sim \mathcal{N}(m,\, \sigma_{\log(y_{i,\mathrm{mean,[ijkl]}})}^2) \\[10pt]

\begin{pmatrix}
\alpha_{0,\mathrm{lab}[i]} \\[4pt]
\alpha_{1,\mathrm{lab}[i]}
\end{pmatrix}
&\sim
\mathcal{N}\!\left(
\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
\Sigma_{\mathrm{lab}}
\right),
\quad
\Sigma_{\mathrm{lab}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{lab}} & \rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} \\[4pt]
\rho_{\mathrm{lab}}\,\sigma_{0,\mathrm{lab}}\,\sigma_{1,\mathrm{lab}} & \sigma^2_{1,\mathrm{lab}}
\end{pmatrix}

\text{, for lab i = 1,} \dots \text{,I} \\[12pt]

\begin{pmatrix}
\alpha_{0,\mathrm{study}[j]} \\[4pt]
\alpha_{1,\mathrm{study}[j]}
\end{pmatrix}
&\sim
\mathcal{N}\!\left(
\begin{pmatrix} 0 \\[4pt] 0 \end{pmatrix},
\Sigma_{\mathrm{study}}
\right),
\quad
\Sigma_{\mathrm{study}} =
\begin{pmatrix}
\sigma^2_{0,\mathrm{study}} & \rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} \\[4pt]
\rho_{\mathrm{study}}\,\sigma_{0,\mathrm{study}}\,\sigma_{1,\mathrm{study}} & \sigma^2_{1,\mathrm{study}}
\end{pmatrix}
\text{, for study j = 1,} \dots \text{,J} \\[12pt]

\alpha_{0,\mathrm{arm}[k]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{arm}})
\text{, for arm k = 1,} \dots \text{,K} \\[12pt]
\alpha_{0,\mathrm{effect}[l]} &\sim \mathcal{N}(0, \sigma^2_{0,\mathrm{effect}})
\text{, for effect l = 1,} \dots \text{,L} \\[12pt]
\end{align*}
$$

@@

The priors for this model were as follows:
$$
\small
\begin{align*}
\beta_0 &\sim \text{Student-}t(3,\, 5.54,\, 0.80) \\[4pt]
\beta_1 &\sim \text{Student-}t(3,\, 0,\, 5.3) \\[4pt]
\beta_2 &\sim \text{Student-}t(3,\, 0,\, 2.5) \\[12pt]

\text{mean}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 7.28,\, 0.62) \\[4pt]
\text{sd}(\tilde{m}) &\sim \text{Half-student-}t(3,\, 0,\, 5) \\[12pt]

\sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 1.08,\, 1.06) \\[4pt]
\sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\,
\sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 2.5) \\[12pt]

\mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt]
\mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1)
\end{align*}

$$
